On the power of oritatami cotranscriptional folding with unary bead sequence\thanks{This work is supported in part by KAKENHI Grant-in-Aid for Challenging Research (Exploratory) No.~18K19779 and JST Program to Disseminate Tenure Tracking System No.~6F36 granted to S.~Z.~F. and S.~S.}

### 2018-12-06 arity1upper.tex
arity1upper
Assumpation of n removed
Thm and Lemma statements fixed
The original was a bit informal, saying "can yield" or something like
that.

In this section we will argue that unary systems at arity $1$ and delay $2$ and $3$, respectively, cannot fold infinite transcripts deterministically. As we will see, in fact, the length of transcripts deterministically foldable by these systems has an upper bound quadratic in the length $n$ of the seed. The main result is the following theorem, which is a direct consequence of Lemmas~\ref{lem:d2a1}  and \ref{lem:d3a1} which follow.
\begin{theorem}[$\delta\in\{ 2,3\}, \alpha = 1$]\label{thm:d23a1}
	A deterministic unary oritatami system at arity $\alpha = 1$ and delay $\delta \in \{2,3\}$ can fold into a terminal conformation of length at most $3n^2+4n+2$.
\end{theorem}

@@

Let us fix some common starting points for Lemmas~\ref{lem:d2a1} and \ref{lem:d3a1}. Let the point where the first transcript bead was fixed be $p$ and let $n=|\mathrm{seed}|+1$. We will argue about the situation when the first bead is stabilized outside $\hexagon_p^n$ (a hexagon of radius $n$). Let this be the $i$th bead of the transcript. Without loss of generality, we can translate the origin $(0,0)$ to the coordinates of bead $i-1$ (which is still in $\hexagon_p^n$), and we can assume that bead $i$ is fixed at $(1,1)$ (see Fig.~\ref{fig:hexagonOut}).

%\subsubsection{\ref{sec:unary}.1 \ \ $\delta = 2$\\}
\begin{lemma}[$\delta=2, \alpha =1$]\label{lem:d2a1}
	A deterministic unary oritatami system of $\delta=2$ and $\alpha = 1$ can fold into a conformation of length at most $3n^2+4n+2$.
\end{lemma}
\begin{proof}
In the elongation that places bead $i$ at $(1,1)$ there are two possibilities. 
\begin{itemize}
	\item $i$ forms a bond with a bead at $(1,0)$.
	\item  $i$ does not bond to anything and $i+1$ is at $(2,1)$ bonding with a bead at $(2,0)$. If there is no bead at $(1,0)$, then placing $i$ at $(1,0)$ instead of $(1,1)$ results in the same number of bonds, leading to nondeterminism. Therefore, there has to be a bead at $(1,0)$ and it is inactive, otherwise it would bond to $i$. This is analogous to case 1. below, with the only difference being that the bond between (1,1) and (1,0) is missing.%as in Fig.~\ref{fig:hexagonOut1}.
\end{itemize}

@@

\begin{lemma}[$\delta=3, \alpha =1$]\label{lem:d3a1}
	A deterministic unary oritatami system of $\delta=3$ and $\alpha = 1$ can fold into a terminal conformation of length at most $3n^2+4n+2$.
\end{lemma}
\begin{proof}
We will argue similarly to the $\delta=2$ case: when we stabilize beads outside $\hexagon_p^n$, we can only do so at points right next to one of the sides and even there only until we reach a corner. This yields the upper bound immediately.

>>>>

In this section we will argue that unary systems at arity $1$ and delay $2$ and $3$, respectively, cannot fold infinite transcripts deterministically. As we will see, in fact, the length of transcripts deterministically foldable by these systems has an upper bound quadratic in the length $n$ of the seed. The main result is the following theorem, which is a direct consequence of Lemmas~\ref{lem:d2a1}  and \ref{lem:d3a1} which follow.
\begin{theorem}[$\delta\in\{ 2,3\}, \alpha = 1$]\label{thm:d23a1}
	The terminal conformation of a deterministic unary oritatami system at arity $\alpha = 1$ and delay $\delta \in \{2,3\}$ is of length at most $3n^2+4n+2$.
\end{theorem}

@@

Let us fix some common starting points for Lemmas~\ref{lem:d2a1} and \ref{lem:d3a1}. Let the point where the first transcript bead was fixed be $p$. We will argue about the situation when the first bead is stabilized outside $\hexagon_p^n$ (a hexagon of radius $n$). Let this be the $i$th bead of the transcript. Without loss of generality, we can translate the origin $(0,0)$ to the coordinates of bead $i-1$ (which is still in $\hexagon_p^n$), and we can assume that bead $i$ is fixed at $(1,1)$ (see Fig.~\ref{fig:hexagonOut}).

%\subsubsection{\ref{sec:unary}.1 \ \ $\delta = 2$\\}
\begin{lemma}[$\delta=2, \alpha =1$]\label{lem:d2a1}
	The terminal conformation of a deterministic unary oritatami system of $\delta=2$ and $\alpha = 1$ is of length at most $3n^2+4n+2$.
\end{lemma}
\begin{proof}
In the elongation that places bead $i$ at $(1,1)$ there are two possibilities. 
\begin{itemize}
	\item $i$ forms a bond with a bead at $(1,0)$.
	\item  $i$ does not bond to anything and $i+1$ is at $(2,1)$ bonding with a bead at $(2,0)$. If there is no bead at $(1,0)$, then placing $i$ at $(1,0)$ instead of $(1,1)$ results in the same number of bonds, leading to nondeterminism. Therefore, there has to be a bead at $(1,0)$ and it is inactive, otherwise it would bond to $i$. This is analogous to case 1. below, with the only difference being that the bond between (1,1) and (1,0) is missing.%as in Fig.~\ref{fig:hexagonOut1}.
\end{itemize}

@@

\begin{lemma}[$\delta=3, \alpha =1$]\label{lem:d3a1}
	The terminal conformation of a deterministic unary oritatami system of $\delta=3$ and $\alpha = 1$ is of length at most $3n^2+4n+2$.
\end{lemma}
\begin{proof}
We will argue similarly to the $\delta=2$ case: when we stabilize beads outside $\hexagon_p^n$, we can only do so at points right next to one of the sides and even there only until we reach a corner. This yields the upper bound immediately.
